---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

First of all, we're going to load the R-package `ggplot2` for creating nice-looking plots: 
```{r echo=TRUE}
require(ggplot2)
```


## Loading and preprocessing the data
The data is available in the 'activity.zip' file. We're going to read it in directly from the zipped archive:
```{r echo=TRUE}
        data <- read.csv(unz("activity.zip","activity.csv"))
```

Let's see the data formats, using `str`:
```{r echo=TRUE}
        str(data)
```

Since the date column is stored as a factor, let's create column 'date2' in the `POSIXlt` format for later use:
```{r echo=TRUE}
        data$date2 <- strptime(data$date, "%Y-%m-%d")
        str(data)
```

## What is mean total number of steps taken per day?

First, we'll calculate the total number of steps taken per day and print the result in a scatterplot:
```{r echo=TRUE}
        dailysum <- tapply(data$steps, data$date, FUN=sum)
        plot(dailysum, main ="Total steps per day", xlab="Days", ylab="Steps")
```

Then we're going to make a simple histogram plot:
```{r echo=TRUE} 
         ggplot() + aes(dailysum)+ geom_histogram(binwidth=500, colour="black", fill="steelblue", na.rm=TRUE) + ggtitle("Histogram of Total Daily Steps") + xlab("Total daily steps") + ylab("Count (days)")
```

The  mean and median of the total number of steps taken per day are:
```{r echo=TRUE}
        mean <- as.integer(mean(dailysum, na.rm=TRUE))
        median <- as.integer(median(dailysum, na.rm=TRUE))
```
- Median: `r median`
- Mean: `r mean`

_Note the `as.integer`, which I used to prevent double precision that would provide no valuable information._

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
